---
title: "ASB Visualiser"
author: "Terry Fenton"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import ASB CSV files

Below is the function we shall use to read in the CSV's. The output is a list of dataframes.

```{r}
csvReader <- function(dir, dirFiles) {
  
  dfList <- list()
  
  for (i in 1:length(dirFiles)) {
    # Dataframe Labels
    headers <- as.character(read.csv(paste0(dir, dirFiles[i]), 
                                   skip = 6, nrow = 1, header = FALSE)[1,])
    # Dataframe itself
    df <- read.csv(paste0(dir, dirFiles[i]), skip = 8, header = FALSE)
    colnames(df) <- headers
    
    # Add to list of Dataframe
    dfList[[i]] <- df
  }
  
  dfList
  
}
```

Below is where the details are put through.

```{r}
# dir is name of the folder that holds your csv files. This should exist in your 
# directory.
dir <- "dataframes/"
dirFiles <- list.files(dir)
dfList <- csvReader(dir, dirFiles)

head(dfList[[1]])
```

## Merge dataframes

In this process, we want to merge all the dataframes (which represent individual accounts) into one. To do this cleanly we will also have to eliminate all rows which moves funds between accounts.

```{r}
mergeDF <- function(dfList) {
  
  # Input checks
  if (length(dfList) == 0) {
    stop("list length 0")
  }
  
  # Create Single Df
  finDF <- dfList[[1]]
  for (i in 2:length(dfList)) {
    # Row bind the two dataframes
    finDF <- rbind(finDF, dfList[[i]])
  }
  
  finDF
  
}

```

Process dfList into a single dataframe.

```{r}
bigDF <- mergeDF(dfList)

# Modify headers for cleanliness
newHeaders <- gsub(" ", "_", colnames(bigDF))
colnames(bigDF) <- newHeaders
```

## Explore Data

### Remove internal account transactions

```{r}
# TFR IN and TFR OUT balance change
amountTFR <- bigDF$Amount[bigDF$Tran_Type %in% c("TFR IN", "TFR OUT")]

# all.equal returns true if the sum is equivalent to 0
if (all.equal(sum(amountTFR), 0) == FALSE) {
  stop("Internal Account Transfers do not sum to 0, 
       provided datasets may be incomplete")
}

# Remove TFR IN and TFR OUT
dfNoTFR <- bigDF[!(bigDF$Tran_Type %in% c("TFR IN", "TFR OUT")), ]

# Account Change relative to prior day
changesDf <- aggregate(Amount ~ Date, data = dfNoTFR, FUN = sum)

# Fill in gaps (This ended up becoming more complicated than I thought)
periodCoverage <- seq(as.Date(changesDf$Date[1], tryFormats = "%Y/%m/%d"),
                      as.Date(changesDf$Date[dim(changesDf)[1]], 
                              tryFormats = "%Y/%m/%d"),
                      by = "days") |>
  gsub(pattern = "-", replacement = "/")

zeroLots <- c(0) * length(periodCoverage)

# Setup final df with every day in period recorded
accChangeDf <- data.frame(Date = periodCoverage, Amount = zeroLots)

for (date in changesDf$Date) {
  accChangeDf$Amount[accChangeDf$Date == date] <- changesDf$Amount[changesDf$Date == date]
}

```

### Visualise transaction change over time

```{r}
# Date timestamp
monthDay <- accChangeDf$Date |> substr(6, 10)
dateSpan <- paste(monthDay[1], "to", monthDay[length(monthDay)])

# Overall change
overallChange <- round(sum(accChangeDf$Amount), 2)

# Basics
plot.new()
plot.window(range(1:dim(accChangeDf)[1]), range(accChangeDf$Amount))
par(cex = 0.7, omi = c(0, 0, 0, 1))

# Plot Details
points(factor(accChangeDf$Date), accChangeDf$Amount)
abline(0, 0, col = "darkred")
segments(1:dim(accChangeDf)[1], c(0)*dim(accChangeDf)[1],
         1:dim(accChangeDf)[1], accChangeDf$Amount)
axis(1, 1:dim(accChangeDf)[1], monthDay)
axis(2)
box()

# Labeling
condCol <- ifelse(overallChange >= 0, "darkgreen", "darkred")

title(paste("Account Change per Day from", dateSpan, "(mm/dd)"),  
      xlab = "Date", ylab = "Dollar change in NZD ($)")

mtext("Cumulative Diff:", 4, outer = TRUE, las = 1, cex = 0.8)

mtext(paste0("\n\n+",overallChange), 4, outer = TRUE, las = 1, 
      cex = 0.8, col = condCol)

```

Lets turn the above visualisation into a function.

```{r}

relativeChange <- function(noTransDf, period = "Day") {
  # Specified df
  accChangeDf <- aggregate(Amount ~ Date, data = dfNoTFR, FUN = sum)
  
  # Date timestamp
  if (period %in% c("day", "Day", "DAY")) {
    monthDay <- accChangeDf$Date |> substr(6, 10)
    dateSpan <- paste(monthDay[1], "to", monthDay[length(monthDay)])
  } else if (period %in% c("month", "Month", "MONTH")) {
    monthDay <- accChangeDf$Date |> substr(1, 7)
    dateSpan <- paste(monthDay[1], "to", monthDay[length(monthDay)])
    accChangeDf$Date <- monthDay
    accChangeDf <- aggregate(Amount ~ Date, data = accChangeDf, FUN = sum)
  } else {
    stop("Period incorrectly specified. Try 'day' or 'month'")
  }
  
  
  # Overall change
  overallChange <- round(sum(accChangeDf$Amount), 2)
  
  # Basics
  plot.new()
  plot.window(range(1:dim(accChangeDf)[1]), range(accChangeDf$Amount))
  par(cex = 0.7, omi = c(0, 0, 0, 1))
  
  # Plot Details
  points(factor(accChangeDf$Date), accChangeDf$Amount)
  abline(0, 0, col = "darkred")
  segments(1:dim(accChangeDf)[1], c(0)*dim(accChangeDf)[1],
           1:dim(accChangeDf)[1], accChangeDf$Amount)
  axis(1, 1:dim(accChangeDf)[1], unique(monthDay))
  axis(2)
  box()
  
  # Labeling
  condCol <- ifelse(overallChange >= 0, "darkgreen", "darkred")
  specifier <- ifelse(period %in% c("day", "Day", "DAY"), "(mm/dd)", "(yyyy/mm)")
  prefix <- ifelse(period %in% c("day", "Day", "DAY"), "Daily", "Monthly")
  
  title(paste(prefix, "Balance Change from", dateSpan, specifier),  
        xlab = "Date", ylab = "Dollar change in NZD ($)")
  
  mtext("Cumulative Diff:", 4, outer = TRUE, las = 1, cex = 0.8)
  
  mtext(paste0("\n\n+",overallChange), 4, outer = TRUE, las = 1, 
        cex = 0.8, col = condCol)
}

```

Test Function

```{r}
relativeChange(dfNoTFR, "Day")
relativeChange(dfNoTFR, "Month")

```

### Gather account net change

Below are functions which will forfill our objectives

```{r}
# Read CSV's to find total amount on the last day
csvAvailBal <- function(dir, dirFiles) {
  
  balances <- as.numeric()
  
  for (i in 1:length(dirFiles)) {
    # Extract available balance
    availBalStr <- as.character(read.csv(paste0(dir, dirFiles[i]), 
                            skip = 4, nrow = 1, header = FALSE))
    
    availBal <- as.numeric(strsplit(availBalStr, " ", fixed = TRUE)[[1]][4])
    
    balances[i] <- availBal
  }
  balances
}

# Find cumulative balance for previous days.

priorBalance <- function(date, finalSum, givenDF) {
  # Calculate day difference
  priorDate <- as.Date(gsub("/", "-", date))
  finalDate <- as.Date(gsub("/", "-", givenDF$Date[dim(givenDF)[1]]))
  daysPrior <- finalDate - priorDate
  
  # Check if same day
  if (daysPrior == 0) {
    return(finalSum)
  }
  
  # Traverse Backwards through each day
  reversedAmounts <- givenDF$Amount[dim(givenDF)[1]:1]
  currentSum <- finalSum
  
  for (i in 1:daysPrior) {
    ##as.Date(-i, origin = givenDF$Date[dim(givenDF)[1]])
    currentSum <- currentSum - reversedAmounts[i]
  }
  
  currentSum
}
```

Generate overall account balance.

```{r}
# Run function and grab final Balance
availBal <- csvAvailBal(dir, dirFiles)
cumulativeBal <- sum(availBal)

# Check balances for all previous dates in df
balances <- as.numeric()
for (i in 1:dim(accChangeDf)[1]) {
  given_date <- accChangeDf$Date[i]
  balances[i] <- priorBalance(given_date, cumulativeBal, accChangeDf)
}

balanceDf <- data.frame(Date = accChangeDf$Date, Balance = balances)
tail(balanceDf)
```

### Visualise Account Balance over period alongside rel change

```{r}

plotOverview <- function(balDf, transDf) {
  # Joint layout
  layout(matrix(c(1, 1, 2, 2), nrow = 2, byrow = TRUE))
  
  # Alter balances to be averaged by month
  balDf$Date <- balDf$Date |> substr(1, 7)
  balDf <- aggregate(Balance ~ Date, data = balDf, FUN = function(x) {x[length(x)]})
  dateSpan <- paste(balDf$Date[1], "to", balDf$Date[dim(balDf)[1]])
  
  # Graph of average balances by Month
  par(cex = 0.7, omi = c(0, 0, 0, 1))
  plot.new()
  plot.window(range(1:dim(balDf)[1]), range(balDf$Balance))
  points(factor(balDf$Date), balDf$Balance, pch = 20)
  lines(factor(balDf$Date), balDf$Balance)
  axis(2)
  box()
  title(paste("Monthly Balance Over", dateSpan, "(yyyy/mm)"), ylab = "Balance Amount ($)")
  
  # Graph of relative changes by Month
  relativeChange(transDf, "Month")
}

```

Plot everything

```{r}
plotOverview(balanceDf, dfNoTFR)
```

## Transaction Classifier

Now with visual changes determined, understanding where that money is going could be incredibly beneficial for understanding the context behind these transactions.




