---
title: "ASB Visualiser"
author: "Terry Fenton"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import ASB CSV files

Below is the function we shall use to read in the CSV's. The output is a list of dataframes.

```{r}
csvReader <- function(dir, dirFiles) {
  
  dfList <- list()
  
  for (i in 1:length(dirFiles)) {
    # Dataframe Labels
    headers <- as.character(read.csv(paste0(dir, dirFiles[i]), 
                                   skip = 6, nrow = 1, header = FALSE)[1,])
    # Dataframe itself
    df <- read.csv(paste0(dir, dirFiles[i]), skip = 8, header = FALSE)
    colnames(df) <- headers
    
    # Add to list of Dataframe
    dfList[[i]] <- df
  }
  
  dfList
  
}
```

Below is where the details are put through.

```{r}
# dir is name of the folder that holds your csv files. This should exist in your 
# directory.
dir <- "dataframes/"
dirFiles <- list.files(dir)
dfList <- csvReader(dir, dirFiles)

# View the list
dfList
```

## Merge dataframes

In this process, we want to merge all the dataframes (which represent individual accounts) into one. To do this cleanly we will also have to eliminate all rows which moves funds between accounts.

```{r}
mergeDF <- function(dfList) {
  
  # Input checks
  if (length(dfList) == 0) {
    stop("list length 0")
  }
  
  # Create Single Df
  finDF <- dfList[[1]]
  for (i in 2:length(dfList)) {
    # Row bind the two dataframes
    finDF <- rbind(finDF, dfList[[i]])
  }
  
  finDF
  
}

```

Process dfList into a single dataframe.

```{r}
bigDF <- mergeDF(dfList)
```

## Explore Data

```{r}
# TFR IN and TFR OUT balance change
amountTFR <- bigDF$Amount[bigDF$`Tran Type` %in% c("TFR IN", "TFR OUT")]

# all.equal returns true if the sum is equivalent to 0
all.equal(sum(amountTFR), 0)

# Remove TFR IN and TFR OUT
```



