---
title: "NB Multinomial Report"
author: "Terry Fenton"
date: "`r Sys.Date()`"
output: html_document
---

# Scores

```{r}
# Accuracy: (TP + TN) / Tot 
trainAC <- sum(trainDF$Classifier == trainPreds) / nrow(trainDF) 
valAC <- sum(valDF$Classifier == valPreds) / nrow(valDF) 
testAC <- sum(testDF$Classifier == testPreds) / nrow(testDF) 

print("Accuracy")
paste("Train:", trainAC)
paste("Validation:", valAC)
paste("Test:", testAC)
```
Overall not a great performance. Let's see where we could improve.

```{r}
# Precision: (TP / (TP + FP))

trainPR <- c()
valPR <- c()
testPR <- c()

for (i in 1:8) {
  trainPR <- c(trainPR, sum(trainDF$Classifier == i & trainPreds == i) / (sum(trainDF$Classifier == i)))
  valPR <- c(valPR, sum(valDF$Classifier == i & valPreds == i) / (sum(valDF$Classifier == i)))
  testPR <- c(testPR, sum(testDF$Classifier == i & testPreds == i) / (sum(testDF$Classifier == i)))
}
precisionTable <- data.frame(classifier = categories[2:9], train = trainPR, validation = valPR, test = testPR)

print("Precision")
precisionTable
```

We are incredibly well suited at predicting Income and Food spending, and are very consistent at predicting expenses 66% of the time across the board. What this says to me is that our regular, recurring expenses have been captured with exception to Expenses (which seems to have a particular expense that is regularly misclassified) and Travel (which for all intensive purposes, should just be AT). 

Health, Education, Investments and Wants are either incredibly variable or scarse based on these scores.

```{r}
# Recall: TP / (TP + FN)

trainRC <- c()
valRC <- c()
testRC <- c()

for (i in 1:8) {
  trainRC <- c(trainRC, sum(trainDF$Classifier == i & trainPreds == i) / 
                 (sum(trainDF$Classifier == i & trainPreds == i) + sum(trainDF$Classifier == i & !(trainPreds == i))))
  valRC <- c(valRC, sum(valDF$Classifier == i & valPreds == i) / 
               (sum(trainDF$Classifier == i & trainPreds == i) + sum(trainDF$Classifier == i & !(trainPreds == i))))
  testRC <- c(testRC, sum(testDF$Classifier == i & testPreds == i) / 
                (sum(trainDF$Classifier == i & trainPreds == i) + sum(trainDF$Classifier == i & !(trainPreds == i))))
}
recallTable <- data.frame(classifier = categories[2:9], train = trainRC, validation = valRC, test = testRC)

print("Recall")
recallTable

```


```{r}
# F1 Score

trainF1 <- c()
valF1 <- c()
testF1 <- c()

for (i in 1:8) {
  trainF1 <- c(trainF1, (2 * trainPR[i] * trainRC[i]) / (trainPR[i] + trainRC[i]))
  valF1 <- c(valF1, (2 * valPR[i] * valRC[i]) / (valPR[i] + valRC[i]))
  testF1 <- c(testF1, (2 * testPR[i] * testRC[i]) / (testPR[i] + testRC[i]))
}

f1Table <- data.frame(classifier = categories[2:9], train = trainF1, validation = valF1, test = testF1)

print("F1 Score")
f1Table

```

Overall, not great. Lets try a different model.



